---
title: "Regression Project"
authors: "Thomas Townsley, John Carr, Matthew Lewis"
date: "December 7th, 2018""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(readr)
```

## R Markdown

This is the markdown file for our final project. 


```{r 1.1}
# 1. Abstract. One parapgraph summary what we set out to learn and what final conclusions were. 
```



```{r 2.1}
# 2. Introduction. Discussion of the research and business in question. 
```


```{r 3.1}

# 3. Overview of the dataset

```


<<<<<<< HEAD
```{r}
=======
```{r 4.1}

>>>>>>> f0e7ed1c33ae989982b136d640b5e992caabb988
# 4. Exploratory analysis

# TIPS from Prof. Stevens: use barplots for categorical variables 
# Use GGalley package to expand ggplot2. Includes histogram and correlation pair plotting functions

library(GGally) #if there's trouble installing the package, go to tools -> global options -> packages -> change CRAN mirror to Oakridge,TN

ggpairs(houses)
houses <- read.csv("DC_properties_clean.csv")

ggplot(houses, aes(x = BATHRM))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Bathrooms", title="Histogram of Bathrooms")

ggplot(houses, aes(x = HF_BATHRM))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Half Bathrooms", title="Histogram of Half Bathrooms")

ggplot(houses, aes(x = ROOMS))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Rooms", title="Histogram of Rooms")

ggplot(houses, aes(x = BEDRM))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Bed Rooms", title="Histogram of Bed Rooms")

ggplot(houses, aes(x = AYB))+
  geom_histogram(color="black", fill="light gray", binwidth = 10)+
  labs(x="Year Built", title="Histogram of Year Built")

ggplot(houses, aes(x = YR_RMDL))+
  geom_histogram(color="black", fill="light gray", binwidth = 10)+
  labs(x="Year Remodelled", title="Histogram of Year Remodelled")

ggplot(houses, aes(x = STORIES))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Stories", title="Histogram of Stories")

ggplot(houses, aes(x = SALEYEAR))+
  geom_histogram(color="black", fill="light gray", binwidth = 5)+
  labs(x="Year Sold", title="Histogram of Year Sold")

ggplot(houses, aes(x = SALEMONTH))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Month Sold", title="Histogram of Month Sold")

ggplot(houses, aes(x = SALEDAY))+
  geom_histogram(color="black", fill="light gray", binwidth = 4)+
  labs(x="Day Sold", title="Histogram of Day Sold")

ggplot(houses, aes(x = GBA))+
  geom_histogram(color="black", fill="light gray", binwidth = 100)+
  labs(x="Square Footage", title="Histogram of Square Footage")



View(houses)
print(levels(houses$CNDTN))
houses$CNDTN <- factor(houses$CNDTN, levels = c("Fair", "Average", "Good", "Very Good", "Excellent"))

#fit <- lm(data = houses, PRICE ~ BATHRM + HF_BATHRM + ROOMS + BEDRM + AYB + YR_RMDL + STORIES + GBA + LANDAREA)
fit <- lm(data = houses, PRICE ~ BATHRM + BEDRM + GBA + LANDAREA + CNDTN + SALEDATE)
summary(fit)




```


```{r 5.1}

# 5. Identification and evaluation of suitable model

```


```{r 6.1}

# 6. Application of the model 

```


```{r 7.1}

# 7. Limitations and assumptions 

```



