---
title: "Regression Project"
authors: "Thomas Townsley, John Carr, Matthew Lewis"
date: "December 7th, 2018""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(readr)
library(GGally)
```

## R Markdown

This is the markdown file for our final project. 


```{r 1.1}
# 1. Abstract. One parapgraph summary what we set out to learn and what final conclusions were. 
```



```{r 2.1}
# 2. Introduction. Discussion of the research and business in question. 

#Our goal is to build a model that can predict the price of a home in the Washington DC area. If we had pertinent data on that home, based on the same data to other homes in the area and their prices, but not the price, we will be able to predict what the price should be. Also, this model would be useful in determining what the price of a home in DC should cost even if there is a price already associated with it. For example, if one were to be in the market for such a home, this model would help the buyer or seller understand what would be a reasonable final price. This would inform the amount a buyer could/should offer, or the price a seller could justifiably list the home at. Data such as land area of the home, number of rooms, year built, etc. will power the predictive algorithym. As we began the discovery of this project, we anticipated that variables such as land area and number of rooms would be the most significant. 
```


```{r 3.1}

# 3. Overview of the dataset

# single observation NOT YET excluded

#Our dataset is titled D.C. Residential Properties and was acquired on kaggle.com. This dataset shows real property information, including most recent sales prices as of July 2018, for properties located in Washington, D.C. There are a number of variables that we beleive would be useful to us in achieving our goal, with "Condition" being the main conditional variable we use. We eliminated some variables that we decided would not be relevant or useful immediately. 

# One way we manipulated the original dataset is by transforming the SALEDATE variable into separate variables for SALEYEAR, SALEMONTH and SALEDAY. We did this because we found little significance to SALEDATE as each was unique and spreadout, and thought there might be more significance in using individual parts of the dates at higher levels. We kept the original SALEDATE column in the dataset, as well.

# Description of variables:
# BATHRM (numeric): Number of full bathrooms
# HF_BATHRM (numeric): Number of half bathrooms (no bathtub or shower)
# ROOMS (numeric): Total number of rooms
# BEDRM (numeric): Number of bedrooms
# AYB (numeric): The earliest time the main portion of the building was built
# YR_RMDL (numeric): Year structure was remodeled
# STORIES (numeric): Number of stories in primary dwelling
# SALEDATE (date): Date of most recent sale
# SALEMONTH (numeric): Column we created for the calendar month of the most recent sale
# SALEDAY (numeric): Column we created for the calendar day of the most recent sale
# SALEYEAR (numeric): Column we created for the calendar year of the most recent sale
# GBA (numeric): Gross building area in square feet
# CNDTN (factor): Condition
# LANDAREA (numeric): Land area of property in square feet
# PRICE (numeric): Price of most recent sale

```


<<<<<<< HEAD
```{r}
=======
```{r 4.1}

# 4. Exploratory analysis

# TIPS from Prof. Stevens: use barplots for categorical variables 
# Use GGalley package to expand ggplot2. Includes histogram and correlation pair plotting functions

 #if there's trouble installing the package, go to tools -> global options -> packages -> change CRAN mirror to Oakridge,TN

houses <- read.csv("DC_properties_clean.csv")

##pair plotting exploratory analysis

#First step: look at all possible correlations minus sale date because it wouldn't allow that many factors in the pairs function ¯\_(ツ)_/¯

int.vectors <- data.frame(houses$HF_BATHRM, houses$BATHRM, houses$ROOMS, houses$BEDRM,houses$AYB, houses$YR_RMDL, houses$PRICE, houses$SALEDAY, houses$SALEMONTH, houses$SALEYEAR, houses$STORIES, houses$LANDAREA, houses$GBA, houses$CNDTN)

View(int.vectors)

ggpairs(int.vectors)

#Observations
# Looks like the variables ROOMS, BATHRM, BEDRM, GBA, STORIES, LANDAREA, SALEYEAR, and PRICE have the highest correlation across the board i.e > |.35|. 

#Next making a pairs plot with just the above variables

int.vectors2 <- data.frame(houses$ROOMS, houses$BATHRM, houses$BEDRM, houses$GBA, houses$LANDAREA, houses$SALEYEAR, houses$PRICE)

ggpairs(int.vectors2)



#Histograms for each variable
ggplot(houses, aes(x = BATHRM))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Bathrooms", title="Histogram of Bathrooms")

ggplot(houses, aes(x = HF_BATHRM))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Half Bathrooms", title="Histogram of Half Bathrooms")

ggplot(houses, aes(x = ROOMS))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Rooms", title="Histogram of Rooms")

ggplot(houses, aes(x = BEDRM))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Bed Rooms", title="Histogram of Bed Rooms")

ggplot(houses, aes(x = AYB))+
  geom_histogram(color="black", fill="light gray", binwidth = 10)+
  labs(x="Year Built", title="Histogram of Year Built")

ggplot(houses, aes(x = YR_RMDL))+
  geom_histogram(color="black", fill="light gray", binwidth = 10)+
  labs(x="Year Remodelled", title="Histogram of Year Remodelled")

ggplot(houses, aes(x = STORIES))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Stories", title="Histogram of Stories")

ggplot(houses, aes(x = SALEYEAR))+
  geom_histogram(color="black", fill="light gray", binwidth = 5)+
  labs(x="Year Sold", title="Histogram of Year Sold")

ggplot(houses, aes(x = SALEMONTH))+
  geom_histogram(color="black", fill="light gray", binwidth = 1)+
  labs(x="Month Sold", title="Histogram of Month Sold")

ggplot(houses, aes(x = SALEDAY))+
  geom_histogram(color="black", fill="light gray", binwidth = 4)+
  labs(x="Day Sold", title="Histogram of Day Sold")

ggplot(houses, aes(x = GBA))+
  geom_histogram(color="black", fill="light gray", binwidth = 100)+
  labs(x="Square Footage", title="Histogram of Square Footage")



View(houses)
print(levels(houses$CNDTN))
houses$CNDTN <- factor(houses$CNDTN, levels = c("Fair", "Average", "Good", "Very Good", "Excellent"))

#fit <- lm(data = houses, PRICE ~ BATHRM + HF_BATHRM + ROOMS + BEDRM + AYB + YR_RMDL + STORIES + GBA + LANDAREA)
fit <- lm(data = houses, PRICE ~ BATHRM + BEDRM + GBA + LANDAREA + CNDTN + SALEDATE)
summary(fit)





```


```{r 5.1}

# 5. Identification and evaluation of suitable model

```


```{r 6.1}

# 6. Application of the model 

```


```{r 7.1}

# 7. Limitations and assumptions 

```



